\section{Reading notes}

This section contains specific references I want to further discuss, or simply notes on references I want to write down.


\subsection{Money creation in the modern economy (M. McLeay, A. Radia, \& R. Thomas)}

After reading \cite{BOE2014}, I felt that there wasn't much substance to the actual article; however, it has led to some other references on money supply and inflation.


\subsection{Will the securitization revolution spread? (R. Feldman)}

In \cite{Feldman1995}, Feldman considers why small business loans have not been securitized to the extent of mortgages, auto loans, and credit card receivables.  This discussion, however, is from 1995 and so this should be more of a jumping-off point than taken as an explanation of the modern situation for small business loan securitization.

Feldman notes that the benefits to securitization are obvious (increased liquidity for lenders, increased loans for borrowers, stable income opportunities for investors), so a natural suspect for the lack of adoption is the difficulty (i.e., cost) in securitization.
\begin{itemize}
    \item Small business loans require detailed information on the management team and the local economy in which the business operates.
    \item Small business loans lack (or, at least they might've in the past) standardized terms.
    \item There is a lack of publicly-available, long-term performance data on small business loans.  I.e., it's a more opaque market when compared to mortgage loans.
\end{itemize}

I find this first point interesting, as it should presumably apply for mortgages and credit card receivables.  I.e., a bank requires detailed income and credit history for a mortgage applicant and a credit card company needs the same to establish a revolving line of credit.  \note{[Consider]:} {\it how does the future income calculation differ between a small business and an individual?}  Feldman notes that this cost/lack of understanding is mostly on the investor side (lenders are already capable of making the loans), who will choose a better-understood investment like mortgage-backed securities when given the option.

Feldman further notes that investors want loss protections when they don't understand the mechanics of the securities they're purchasing.  These loss protections, however, imply that the bank (or, more generally, the lender) still carries the credit risk of the securitized loans.  Maintainin the credit risk requires banks to buy costly insurance against losses on securitized loans.  \note{[Consider]:} {\it how do SBA 7(a) guarantees help with this?}  Moreover, Feldman states that bank regulations consider loans that still carry credit risk to have not been fully sold; i.e., the bank must still meet capital requirements associated with having the securitized loans on its books.  From the bank's point of view, securitization provides liquidity but does not remove the accounting complexity or credit risk associated with the loans.  From the investors point of view, small business loan securities are a complicated--and risky--asset that are not worth the trouble once all the required protections are priced into the return.  It's re-emphasized throughout that the informational burden on the investor is the primary cost preventing broader adoption of securitized small business loans.

\note{[Competition]:} Feldman notes that small business loans are less subject to non-local (lending) competition.  Worth looking into whether this is still the case 30 years later.

\note{[Loan standards]:} it's noted that not all loans are made through the SBA program.  It's worth looking into the size of this non-SBA loan market.


\subsubsection{Summary}

This has been a thought-provoking article that fleshes out the pros and cons associated with small business loan securitization.  Feldman's points here should be contrasted with the current small business loan market.


\subsection{R. Shiller - From efficient markets theory to behavioral finance}

In \cite{Shiller2003}, Shiller puts forth that stock price volatility is at odds with the efficient market hypothesis, which he claims requires stock prices to equal the optimal forecast of the present value of all future dividends from that stock.  In particular, Shiller requires market movements in an ``efficient" market to result from new information that causes market participants to rebalance their portfolios.  Consistent volatility that exceeds informational shortcomings, however, would discredit Shiller's construction of the efficient market hypothesis.

Shiller goes on to compare plots of present value calculations against stock price trends over the past century.  He claims that there is too much volatility in the S\&P to positively support the efficient market hypothesis.  Confusingly, it seems that there is some predictive power in the price of individual stocks about their present value, but that no simple weighting or time-discounting scheme appears to reproduce stock prices.  In particular, no simple scheme of determining the present value of stock divident payouts seems to accurately track boom and bust cycles.

Instead, Shiller transitions to discussing behavioral psychology, where people are shown to exhibit irrational strategies (or heuristics) that would lend themselves to boom-bust cycles in trading markets.  He relates the fervor of a Ponzi scheme to market bubbles, where initial investors get such high returns that new investors are no longer considered with the fundamentals that created the returns.  People simply want in on the action and embrace the risk.

Shiller goes on to discuss the efficient market answer to irrational market participants: that ``smart money'' will move opposite the irrational participants to profit off their mistakes.  An issue Shiller brings up is an apparently-widespread issue that it can be difficult to exert downward pressure on an asset bubble.  He notes the example of 3Com's Palm offering, where Palm shares rose to such heights that the remaining 3Com business had an implied negative value on their overall holdings.  An issue in this bubble was that investors who realized the over-pricing of Palm could not easily short Palm stock due to shortages and high price barriers to put options.  While it's stated that this state of affairs is relatively rare, this sounds a lot like discussion around the housing bubble of the mid-00s.  Even though people had concerns about housing prices, there was no obvious/efficient mechanism for skeptical investors to push prices down.\footnote{Of course, there is the story of {\it The Big Short}, where clever investors made money through credit default swaps with unsuspecting banks.  These investors, however, did not have any actual capacity to sell the underlying homes (and push prices down).}

This paper wraps up with a summary that the stock market certainly reflects information about the companies being bought and sold.  It may be surprising in how efficient it ``normally'' operates, but it is certainly not perfectly efficient.  Shiller advises that economists (or investors, or anyone really) should keep in mind that the market prices are not perfect and that irrational behaviors are a necessary consequence of humans being the buyers and sellers in the market.  I think this paper was most interesting to me as an introduction to how Shiller--and others--value stocks, and think of them as dividend return vehicles.  Any alternative pricing strategy appears to be implicitly ``irrational.''


\note{[Consider]}:
\begin{itemize}
    \item Do I agree that the error is necessarily uncorrelated with the current price/information?
    \item E.g., if a crucial piece of information is missing from the pubilc knowledge, and that piece of information would drastically alter the price, wouldn't the error necessarily depend on the price?
    \item Cast another way: if the error is the actual minus the forecast, then how can this residual not depend on the forecast?
    \item What {\it causes} the uncertainty in $P_t^*$?  Is it purely the correct discount rate for each future dividend?  Presumably it is based off the actual dividends that are eventuallly paid out, rather than being another estimate.
    \item In what universe is the uncertainty in the present value of all future {\it known} dividend payouts bounded (below) by the uncertainty in the market forecast of that present value?  If the fed guaranteed 100 years of interest rates in quarterly intervals, how would the uncertainty in the forecasted present value of future possible dividend payments continue to bound the uncertainty in the present value of actual (known) future dividend payments?
    \item Am I misunderstanding the concept of {\it variance} in this discussion?  Why would the forecast of a random variable ever have a smaller variance (i.e., uncertainty) than the variable itself?
    \item How can it be that per-stock analyses show reasonable price forecasting through the present value, but that this apparently fails on the broader market?  Is this a tail effect where the ``market average'' is skewed by the exceptionally unpredictable stocks?  Do portfolios of stocks somehow create more uncertain forecasts than individual ones?
    \item If boom and bust cycles are removed ({\it if} they can be removed), does the stock price more accuratly reflect present value?  Are boom/bust cycles an aggregate market phenomenon that detracts from an otherwise present-value forecasted stock pricing market?
\end{itemize}


\subsection{K. Cole, J. Helwege, and D. Laster - Stock market valuation indicators: is this time different?}

This paper discusses the state of the stock market following the summer of 1995.  It's noted that the 2.5\% dividend yield in the summer of 1995 was a historical low.  It's further noted that the market-to-book ratio over 3 exceeds the historical average of 1.9.  It starts out by noting that low dividend yields and high market-to-book ratios (stock market capitalization relative to the accounting book value of a company) tend to signal poor future performance.  Counter-arguments to this thinking is allegedly that the era of tax-advantaged stock buybacks and new accounting rules have resulted in an artificial increase in market-to-book ratios and an artificial decrease in dividend yields.  {\it Note: I can't help but notice the echoes of Robert Shiller's critique of the efficient market hypothesis, where he mentioned that new techniques and rationales emerge in a bubble to support ``the new norm'' of eternally-high stock prices and metrics.}  The trio aim to rebuke these claims and sound the alarm on over-priced stocks in this paper.

There's some interesting discussion about share repurchases vs. dividend payouts.  Aside from the obvious tradeoffs (taxes, primarily), it's noted that a lot of the negative connotation with decreased dividend yields or canceled dividends doesn't seem to weigh down share repurchases.  For example, announced share buybacks that don't come to fruition or scaled back share buyback plans don't seem to cause investor alarm.  From this perpspective, the trio consider whether share repurchases are equally valuable to investors, on a dollar-per-dollar basis.  It's noted further that share issuances acts to negate share buybacks.  On this basis, the net share buybacks (share repurchases minus share issuances) is claimed to still highlight low investor return in the 90s despite high stock prices.

The paper concludes by more or less summarizing that, even when share buybacks and new tax rules are taken into account, the current (1995) market appeared over-valued.  Of course, the stock market (as measured via the S\&P 500) would continue roaring until late 2001, where it faultered multiple times until appearing to regain its rapid growth in the 2010s.  The stock market may have been overvalued, but instead of returning to lower level, it repeatedly crashed in the 2000s while trying to grow faster than the fundamentals would support.  Personally, this begs the question of whether market fundamentals ever truly act to drag the price down, or if the market grows as fast as investors will support.

\note{[Consider]}:
\begin{itemize}
    \item The S\&P {\bf 500} (note, not the S\&P 400 used for market cap analysis in the text) dividend yield was a measly 1.17\% in September 2025, an apparent historical low.  I'm guessing from the data that the norm is closer to 2\%, which is still low in the eyes of the text.  A Google search suggests a market-to-book ratio somewhere between 2.5 and 2.9, which is on the higher side but not as inflated as the text describes.  
\end{itemize}


\subsection{K. Case and R. Shiller - Is there a bubble in the housing market?}

\note{[Todo]}: Read \cite{Case2003}


\subsection{L. Zeng, P. Luk - Examining share repurchases and the S\&P buyback indices in the U.S. market}

In \cite{Zeng2020}, Liyu Zeng backtests the S\&P 500 Buyback Index against their other indices.

\note{[Caution]}: The 20 year period in question is likely one of the worst for the S\&P 500, as it includes both the dot com bubble and the 2008 financial crisis, with only a decade surrounding these shortcomings.  The annualized rate of return is only 6.1\% over this period, which suggests to me that it is not representative of the long-term 10\% annualized returns typically seen over 30+ year periods.  The buyback index in question was also mostly backtested data, since the index didn't launch until 2013.  These findings should be re-done on a longer time horizon, possibly in 2030 if the era of buybacks is considered to be post-2000.

The S\&P 500 Buyback Index tracks the 100 companies in the S\&P 500 with the highest buyback ratio, where the buyback ratio is the trailing 12 month amount of share repurchases divided by the total market capitalization at the {\it beginning} of the trailing 12 month period.  Curiously, this is an {\it equally-weighted} index.  This index is apparently adjusted quarterly in January, April, July, and October.

The returns of the S\&P 500 Buyback Index are analyzed over the 20 year period from 1999 to 2019.  The buyback index apparently outperformed the standard S\&P 500 over most of the 20 years, and the buyback index is supposed to have experienced 5.5\% higher returns per year in an annualized basis.

\note{[Yield is king]}: it's not emphasized, but the yield portfolio--the top 100 companies in the S\&P 500 when both dividends and buybacks are used in the trailing 12 month calculation--out-performed the buyback index with lower volatility.  There's an aside that the yield index didn't out-perform the S\&P 500 in as many years as the buyback index, but I'm not sure why the yearly comparison would matter more than the annualized returns over the 20 year period.

Exhibit 6 is an interesting set of data.  Of the five indices compared (buyback, dividend, all-yield, S\&P 500, and {\it equally-weighted} S\&P 500), the standard S\&P 500 performed the worst over the 20 year period.  I would be very curious to see 30 year data, if it's available.  If anything, I think this chart is saying that recent-term yield is a better indicator of future returns than current (or recent) market capitalization.  That feels a bit tautological, but perhaps it's interesting that a one-year run up is enough time to gauge a company's value-returning habits.  The equally-weighted S\&P 500 out-performing the standard index over the 20 year period suggests that there are some large capitalization companies not returning value to shareholders as effectively as other smaller--but still large--capitalization companies.

There's a discussion that the bull cycle momentum catapults the buyback index over the standard index in up months vs. down months.  Exhibit 9 mostly demonstrates that dividends anchor losses against down cycles, and capital gains anchor gains in growth cycles.  The buyback index has the largest upside in growth cycles, but the all-yield index has much better protection against down cycles with similar upside in growth cycles.  Again, this seems to imply that the trailing 12-month shareholder returns are a more valuable weighting metric than purely the market capitalization.  That would be an interesting index to try and maintain\ldots

The discussion of the small-cap weighting of these yield indices (buyback, dividend, and all-yield) really begs for a modern analysis, where the past 5-10 years of the S\&P 500 have been dominated by the top 6-7 tech stocks.  I wonder if the small-cap bias would invert the relation from exhibit 9, where the S\&P 500 was out-perfomed by the other indices that weighted indifferently to market capitalization.

In the small- and mid-cap indices, the buyback index really did outperform the all-yield index.  Notaby, the trend of the equal-weight index outperforming the standard S\&P index also held true.  The small- and mid-cap indices actually seemed to better exemplify the analysis of the S\&P 500 buyback index, where growth-cycles propel the small- and mid-cap buyback indices past the standard index and down cycles don't cause as much value loss.  The all-yield indices both showed negative performance in down cycles vs. the standard (weighted by market capitalization) index, where the S\&P 500 all-yield index still maintained increased returns.  The broader trend nevertheless holds, where dividends anchor returns in down cycles and capital gains propel returns in growth cycles.


\subsection{E. Fama, K. French - A five-factor asset pricing model}

E. Fama and K. French start \cite{Fama2015} with a bit of a history lesson on how people have thought of CAPM and pricing models since the 60s and 90s.  

When introducing the models, it began to sound like the models under discussion were simply adding sufficiently many parameters to capture a majority of the effect.  Funnily enough, this appears to be Fama's preferred interpretation.  They reference \cite{Fama1996} as motivating the intepretation that the portfolios as prescribed are sufficiently distinct and numerous as to span some true, unknown set of market variables that effectively predict the future return.

{\bf Takeaway}:  ``There are patterns in average returns related to Size,
B/M, profitability, and investment.'' I think that quote summarizes the findings, particularly when understood as a response to CAPM, where portfolio diversification/risk was the primary driver of returns.  This is a very long, wordy article with lots of tables and appendices, but I'm not sure if more is to be gleaned there.  My understanding of the article is that there are a handful of market factors driving the majority of returns (i.e., rather than there being an innumerable amount), and that the returns of most portfolios can be understood as arising from a decomposition of that portfolio into these factors.  Whether that's true or not is another matter, but that's how I understood the main thesis of their work.


\subsection{Wikipedia article on the Capital Asset Pricing Model (CAPM)}

To get a baseline for historical discussions of risk vs. return and portfolio pricing, I've been reading through the Capital Asset Pricing Model (CAPM) wikipedia article \cite{CAPM2026}.  My current understanding of the thesis behind CAPM is that the returns of a given portfolio are essentially determined by the market returns for the asset classes in the portfolio, weighted by their risk.\footnote{I'm not saying this is true or correct, but that it's my impression of the CAPM assumptions.}  The subsequent work by Fama and French \cite{Fama2015} (see their earlier work as well) notably provides evidence for contributions outside of CAPM.  Criticisms in the Wikipedia artcle seem to boil down to the lack of a time-constant $\beta$, the immeasurability of {\it expected} market returns for all market participants, and the assumption that the risk profile is totally captured by the mean and variance (e.g., ignoring asymmetric downside risk). 

I won't go much further fleshing out the Wikipedia page, but want to keep the link \cite{CAPM2026} for its references (both within and outside of Wikipedia).


\subsection{Wikipedia article on the Arbitrage Pricing Theory (APT)}

One of the links that piqued my interest on Wikipedia's CAPM page \cite{CAPM2026} is the Arbitrage Pricing Model \cite{APT2026}.  The APT is apparently based on the {\it Law Of One Price} (LOOP), where identical goods sold in different markets (e.g., locations, vendors, etc.) should be priced identically when expressed in the same currency.  The LOOP only holds insofar as taxes, regulations, transaction costs, and all other trade frictions or market manipulations are absent, as it represents the transactional ideal where all arbitrage opportunities have reach a single price equilibrium.

The Arbitrage Pricing Theory (APT) itself, seems like a generic conceptual framework for how returns ought to be related to intrinsic market factors, and {\it risk premiums} associated with those factors.  I'm having some difficulty making sense of the various representations in the article \cite{APT2026}, but will make my best-effort attempt to summarize my understanding in what follows.  Both the Wikipedi article and the original paper \cite{Ross1976} emphasize the multi-dimensional nature of a portfolio of assets, but I'm currently struggling to see the reason and so will start by considering a single asset whose rate of return $r$ is under study.  The APT states that, as far as returns as concerned, the rate of return can be understood as arising from the risk-free rate $r_f$, a number of {\it factors} $f_i$, and an error term $\epsilon$,
\begin{equation}
    r = r_f + \epsilon + \sum_i \beta_i f_i \quad.
\end{equation}
The rate of return $r$, the error $\epsilon$, and the factors $f_i$ are random variables.  The risk-free rate $r_f$ and $\beta_i$ are merely coefficients, where the $\beta_i$ are $\beta$-like in that they denote the sensitivity of the asset to the $i$th factor, $f_i$.  The major consequence of this framework is that the expected return is given by the risk free rate $r_f$, the $\beta$-like $\beta_i$ coefficients, and the {\it risk premiums} $\mu_i$ associated with the factors $f_i$.  I.e.,
\begin{equation}
    \langle r \rangle = r_f + \sum_i \beta_i \mu_i \quad ,
\end{equation}
where we presume $\langle \epsilon \rangle = 0$ for the error term.

The main distinction of the APT relative to the CAPM that I currently see is that it relaxes the prescriptive power of CAPM to accommodate a multi-factor explanation for asset prices.  My current understanding is that this kind of framework underlies the modern multi-factor models like the 5-factor model \cite{Fama2015}.  For example, the framework doesn't say what the underlying factors are, aside from suggestively naming the resulting coefficients {\it risk premiums}.  All that we know is there are a linear combination of factors, which increase the return of an asset relative to a reference risk-free rate.  


\subsection{S. Ross - The arbitrage theory of capital asset pricing}

\note{[Todo]}: read \cite{Ross1976} after above\ldots  See also the working paper \cite{Ross1971}, where the rigorous framework was apparently fleshed out.